---
title: "Susceptible Reconstruction With Noise"
author: "Kaitlyn Stocker"
date: "2/7/2017"
output: html_document
---
```{r, include=FALSE}
library(rstan)
library(ggplot2)
```

```{r, include=FALSE}
#Initial Conditions
n.now <- 763000
s.now <- 762995
i.now <- n.now-s.now
r.now <- 0
beta <- 8
birthrate <-.0005 #rgamma(1, shape=2, scale=.000215)
b.now <- 0

#Small Sample Size Initials 
n.small <- 76300
s.small <- 76295
i.small <- n.small - s.small

n.smaller <- 7630
s.smaller <- 7625
i.smaller <- n.small - s.small

#Noise Conditions
s.noise.low <- rnorm(1,0,n.now/1000)
s.noise.high <- rnorm(1,0,n.now/100)
s.noise.mid <- rnorm(1,0,n.now/500)

i.noise.low <- abs(rnorm(1,1,.005))
i.noise.mid <- abs(rnorm(1,1,.01))
i.noise.high <- abs(rnorm(1,1,.05))
  

#Stan Code
Stan_reconstructed <- as.character("
data {
	int<lower=1> N; // time steps
  real P[N]; //population size
  real time_added[N];
	int <lower=0> I[N]; // infecteds 
	int  Z[N]; // Zt 
}

parameters {
  real <lower=P[1]/100, upper= P[N]> Sbar;
	real<lower=0, upper=5> beta;
  real<lower=0, upper=5> gamma;
}

model {
	//priors
  Sbar ~ normal(P[1]/2, P[1]/10);
	beta ~ gamma(3, 1);
  gamma ~ gamma(2,1);

	//likelihood
	for (n in 2:N){
		I[n] ~ neg_binomial_2(beta*I[n-1]*(Z[n-1]+Sbar)/P[n-1], I[n-1]);
    time_added[n]~exponential_log(gamma);
  }
}    ")


Stan_sim <- as.character("
                            data {
	int<lower=1> N; // time steps
  real P[N]; //population size
  real time_added[N];
	int <lower=0> I[N];
	int <lower=0> S[N];
}


parameters {
	real<lower=0, upper=5> beta;
  real<lower=0, upper=5> gamma;

}

model {
	//priors
	beta ~ gamma(3, 1);
  gamma ~ gamma(2,1);
	//likelihood
	for (n in 2:N){
		I[n] ~ neg_binomial_2(beta*I[n-1]*S[n-1]/P[n-1], I[n-1]);
    time_added[n]~exponential_log(gamma);
  }
}     ")
```

Small Sample Size (N_initial = 76,300)
```{r, echo=FALSE}
#Simulate Epidemic Dynamics 
Sim_small_ss <- function(beta, n.now, i.now, s.now, r.now, b.now) {

results <- as.data.frame(matrix(c(i.now, s.now, r.now, b.now, n.now), nrow=1))
names(results) <- c("I", "S", "R", "B", "N")
  while(s.now*i.now >0){
    i.next <- rnbinom(n=1, mu=beta*i.now*s.now/n.now, size=i.now) 
    b.now <- rgamma(1, shape=2, scale=.000215)*n.now
    s.now <- s.now - i.next + b.now
    r.now <- r.now + i.now
    n.now <- n.now + b.now 
    i.now <- i.next
    results <- rbind(results, c(i.now, s.now, r.now, b.now, n.now))
  }
return(results)
}

simdata_small_ss <- Sim_small_ss(beta=beta, n.now=n.small, i.now=i.small, s.now=s.small, r.now=r.now, b.now=b.now)

#Pull out Simulation Data and Create Time Vector 
I <- simdata_small_ss$I
B <- simdata_small_ss$B
S_sim <- simdata_small_ss$S
n <- simdata_small_ss$N
R <- simdata_small_ss$R
T<- cumsum(rexp(length(I), rate=(1/2.2)))
time_added<-0
for(i in 2:(length(T))){
  time_added[i] <- T[i] - T[i-1]
}

population_check <- 0
for(i in 1:length(I)){
population_check[i] <- I[i] + S_sim[i] + R[i] -n[i]
}

population_check

#Plot Simulation
ggplot(simdata_small_ss, aes(x=T)) +
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=I, col="I"))+
  geom_line(aes(y=R, col="R"))+
  labs(x="Time Steps", y = "Number of Individuals", title = "Simulated Dynamics, Population Size 76,300")

#Reconstructing Z dynamics 
cumbirths <- cumsum(B)
cuminfecteds <- cumsum(I)
cumulativeinfo <- as.data.frame(matrix(c(cumbirths, cuminfecteds), nrow=length(cuminfecteds)))
names(cumulativeinfo) <- c("cumbirths", "cuminfecteds")

regression <- lm(cumbirths ~ 1, offset=(cuminfecteds))

Z <- as.vector(resid(regression))
```
```{r, include=FALSE}
#Run Bayes
data_reconstructed <- list(N=length(I), P=as.integer(n), I=as.integer(I), Z=as.integer(Z), time_added=time_added)

bayes_reconstructed <- stan(model_code = Stan_reconstructed, data=data_reconstructed, iter=500000, chains=4)

#Run Bayes for Simulation
data_sim <- list(N=length(I), P=as.integer(n), I=as.integer(I), S=as.integer(S_sim), time_added=time_added)

bayes_sim <- stan(model_code = Stan_sim, data=data_sim, iter=500000, chains=4)
```
```{r, echo=FALSE}
#Plot S_reconstructed
Sbar <- extract(bayes_reconstructed, 'Sbar')
Sbar <- unlist(Sbar, use.names=FALSE)
lowlim_Sbar <- quantile(Sbar, probs=.025, names=FALSE)
uplim_Sbar <- quantile(Sbar, probs=.975, names=FALSE)

S_lower <- lowlim_Sbar + Z
S_upper <- uplim_Sbar + Z

ggplot(simdata_small_ss, aes(x=T))+
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=S_lower, col="S_lower"))+
  geom_line(aes(y=S_upper, col="S_upper"))+
  labs(x="time", y="Number of Susceptibles", title="Reconstructed vs. Simulated Susceptible Dynamics")
```
```{r}
bayes_sim
bayes_reconstructed
```

Small Sample Size (N_initial = 7,630)
```{r, echo=FALSE}
simdata_smaller_ss <- Sim_small_ss(beta=beta, n.now=n.smaller, i.now=i.smaller, s.now=s.smaller, r.now=r.now, b.now=b.now)

#Pull out Simulation Data and Create Time Vector 
I <- simdata_smaller_ss$I
B <- simdata_smaller_ss$B
S_sim <- simdata_smaller_ss$S
n <- simdata_smaller_ss$N
T<- cumsum(rexp(length(I), rate=(1/2.2)))
time_added<-0
for(i in 2:(length(T))){
  time_added[i] <- T[i] - T[i-1]
}

#Plot Simulation
ggplot(simdata_smaller_ss, aes(x=T)) +
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=I, col="I"))+
  geom_line(aes(y=R, col="R"))+
  labs(x="Time Steps", y = "Number of Individuals", title = "Simulated Dynamics, Population Size 7,630")

#Reconstructing Z dynamics 
cumbirths <- cumsum(B)
cuminfecteds <- cumsum(I)
cumulativeinfo <- as.data.frame(matrix(c(cumbirths, cuminfecteds), nrow=length(cuminfecteds)))
names(cumulativeinfo) <- c("cumbirths", "cuminfecteds")

regression <- lm(cumbirths ~ 1, offset=(cuminfecteds))

Z <- as.vector(resid(regression))
```
```{r, include=FALSE}
#Run Bayes
data_reconstructed <- list(N=length(I), P=as.integer(n), I=as.integer(I), Z=as.integer(Z), time_added=time_added)

bayes_reconstructed <- stan(model_code = Stan_reconstructed, data=data_reconstructed, iter=500000, chains=4)

#Run Bayes for Simulation
data_sim <- list(N=length(I), P=as.integer(n), I=as.integer(I), S=as.integer(S_sim), time_added=time_added)

bayes_sim <- stan(model_code = Stan_sim, data=data_sim, iter=500000, chains=4)
```
```{r, echo=FALSE}
#Plot S_reconstructed
Sbar <- extract(bayes_reconstructed, 'Sbar')
Sbar <- unlist(Sbar, use.names=FALSE)
lowlim_Sbar <- quantile(Sbar, probs=.025, names=FALSE)
uplim_Sbar <- quantile(Sbar, probs=.975, names=FALSE)

S_lower <- lowlim_Sbar + Z
S_upper <- uplim_Sbar + Z

ggplot(simdata_smaller_ss, aes(x=T))+
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=S_lower, col="S_lower"))+
  geom_line(aes(y=S_upper, col="S_upper"))+
  labs(x="time", y="Number of Susceptibles", title="Reconstructed vs. Simulated Susceptible Dynamics")
```
```{r}
bayes_sim
bayes_reconstructed
```


Adding Noise to the Susceptible Class: Low Noise 
```{r, echo=FALSE}
#Simulate Epidemic Dynamics 
Sim_snoise <- function(beta, n.now, i.now, s.now, r.now, b.now, s.noise) {

results <- as.data.frame(matrix(c(i.now, s.now, r.now, b.now, n.now), nrow=1))
names(results) <- c("I", "S", "R", "B", "N")
  while(s.now*i.now >0){
    i.next <- rnbinom(n=1, mu=beta*i.now*s.now/n.now, size=i.now) 
    s.now <- s.now - i.next + b.now + s.noise
    r.now <- r.now + i.now
    b.now <- rgamma(1, shape=2, scale=.000215)*n.now
    n.now <- n.now + b.now 
    i.now <- i.next
    results <- rbind(results, c(i.now, s.now, r.now, b.now, n.now))
  }
return(results)
}

s.noise.low <- rnorm(1,0,n.now/1000)
s.noise.high <- rnorm(1,0,n.now/100)
s.noise.mid <- rnorm(1,0,n.now/500)

simdata_snoise_low <- Sim_snoise(beta=beta, n.now=n.now, i.now=i.now, s.now=s.now, r.now=r.now, b.now=b.now, s.noise=s.noise.low)

#Pull out Simulation Data and Create Time Vector 
I <- simdata_snoise_low$I
B <- simdata_snoise_low$B
S_sim <- simdata_snoise_low$S
n <- simdata_snoise_low$N
T<- cumsum(rexp(length(I), rate=(1/2.2)))
time_added<-0
for(i in 2:(length(T))){
  time_added[i] <- T[i] - T[i-1]
}

#Plot Simulation
ggplot(simdata_snoise_low, aes(x=T)) +
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=I, col="I"))+
  geom_line(aes(y=R, col="R"))+
  labs(x="Time Steps", y = "Number of Individuals", title = "Simulated Dynamics with Noise in Susceptible Class, where noise~N(0,n[t-1]/1000)")

#Reconstructing Z dynamics 
cumbirths <- cumsum(B)
cuminfecteds <- cumsum(I)
cumulativeinfo <- as.data.frame(matrix(c(cumbirths, cuminfecteds), nrow=length(cuminfecteds)))
names(cumulativeinfo) <- c("cumbirths", "cuminfecteds")

regression <- lm(cumbirths ~ 1, offset=(cuminfecteds))

Z <- as.vector(resid(regression))

```
```{r, include=FALSE}
#Run Bayes
data_reconstructed <- list(N=length(I), P=as.integer(n), I=as.integer(I), Z=as.integer(Z), time_added=time_added)

bayes_reconstructed <- stan(model_code = Stan_reconstructed, data=data_reconstructed, iter=500000, chains=4)

#Run Bayes for Simulation
data_sim <- list(N=length(I), P=as.integer(n), I=as.integer(I), S=as.integer(S_sim), time_added=time_added)

bayes_sim <- stan(model_code = Stan_sim, data=data_sim, iter=500000, chains=4)
```
```{r, echo=FALSE}
#Plot S_reconstructed
Sbar <- extract(bayes_reconstructed, 'Sbar')
Sbar <- unlist(Sbar, use.names=FALSE)
lowlim_Sbar <- quantile(Sbar, probs=.025, names=FALSE)
uplim_Sbar <- quantile(Sbar, probs=.975, names=FALSE)

S_lower <- lowlim_Sbar + Z
S_upper <- uplim_Sbar + Z

ggplot(simdata_snoise_low, aes(x=T))+
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=S_lower, col="S_lower"))+
  geom_line(aes(y=S_upper, col="S_upper"))+
  labs(x="time", y="Number of Susceptibles", title="Reconstructed vs. Simulated Susceptible Dynamics")
```
```{r}
bayes_sim
bayes_reconstructed
```

Adding Noise to the Susceptible Class: High Noise
```{r, echo=FALSE}
simdata_snoise_high <- Sim_snoise(beta=beta, n.now=n.now, i.now=i.now, s.now=s.now, r.now=r.now, b.now=b.now, s.noise=s.noise.high)

#Pull out Simulation Data and Create Time Vector 
I <- simdata_snoise_high$I
B <- simdata_snoise_high$B
S_sim <- simdata_snoise_high$S
n <- simdata_snoise_high$N
T<- cumsum(rexp(length(I), rate=(1/2.2)))
time_added<-0
for(i in 2:(length(T))){
  time_added[i] <- T[i] - T[i-1]
}

#Plot Simulation
ggplot(simdata_snoise_high, aes(x=T)) +
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=I, col="I"))+
  geom_line(aes(y=R, col="R"))+
  labs(x="Time Steps", y = "Number of Individuals", title = "Simulated Dynamics with Noise in Susceptible Class, where noise~N(0,n[t-1]/100)")

#Reconstructing Z dynamics 
cumbirths <- cumsum(B)
cuminfecteds <- cumsum(I)
cumulativeinfo <- as.data.frame(matrix(c(cumbirths, cuminfecteds), nrow=length(cuminfecteds)))
names(cumulativeinfo) <- c("cumbirths", "cuminfecteds")

regression <- lm(cumbirths ~ 1, offset=(cuminfecteds))

Z <- as.vector(resid(regression))
```
```{r, include=FALSE}
#Run Bayes
data_reconstructed <- list(N=length(I), P=as.integer(n), I=as.integer(I), Z=as.integer(Z), time_added=time_added)

bayes_reconstructed <- stan(model_code = Stan_reconstructed, data=data_reconstructed, iter=500000, chains=4)

#Run Bayes for Simulation
data_sim <- list(N=length(I), P=as.integer(n), I=as.integer(I), S=as.integer(S_sim), time_added=time_added)

bayes_sim <- stan(model_code = Stan_sim, data=data_sim, iter=500000, chains=4)
```
```{r, echo=FALSE}
#Plot S_reconstructed
Sbar <- extract(bayes_reconstructed, 'Sbar')
Sbar <- unlist(Sbar, use.names=FALSE)
lowlim_Sbar <- quantile(Sbar, probs=.025, names=FALSE)
uplim_Sbar <- quantile(Sbar, probs=.975, names=FALSE)

S_lower <- lowlim_Sbar + Z
S_upper <- uplim_Sbar + Z

ggplot(simdata_snoise_high, aes(x=T))+
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=S_lower, col="S_lower"))+
  geom_line(aes(y=S_upper, col="S_upper"))+
  labs(x="time", y="Number of Susceptibles", title="Reconstructed vs. Simulated Susceptible Dynamics")
```
```{r}
bayes_sim
bayes_reconstructed
```

Adding Noise to the Susceptible Class: Mid Noise 
```{r, echo=FALSE}
simdata_snoise_mid <- Sim_snoise(beta=beta, n.now=n.now, i.now=i.now, s.now=s.now, r.now=r.now, b.now=b.now, s.noise=s.noise.mid)

#Pull out Simulation Data and Create Time Vector 
I <- simdata_snoise_mid$I
B <- simdata_snoise_mid$B
S_sim <- simdata_snoise_mid$S
n <- simdata_snoise_mid$N
T<- cumsum(rexp(length(I), rate=(1/2.2)))
time_added<-0
for(i in 2:(length(T))){
  time_added[i] <- T[i] - T[i-1]
}

#Plot Simulation
ggplot(simdata_snoise_mid, aes(x=T)) +
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=I, col="I"))+
  geom_line(aes(y=R, col="R"))+
  labs(x="Time Steps", y = "Number of Individuals", title = "Simulated Dynamics with Noise in Susceptible Class, where noise~N(0,n[t-1]/500)")

#Reconstructing Z dynamics 
cumbirths <- cumsum(B)
cuminfecteds <- cumsum(I)
cumulativeinfo <- as.data.frame(matrix(c(cumbirths, cuminfecteds), nrow=length(cuminfecteds)))
names(cumulativeinfo) <- c("cumbirths", "cuminfecteds")

regression <- lm(cumbirths ~ 1, offset=(cuminfecteds))

Z <- as.vector(resid(regression))

```
```{r, include=FALSE}
#Run Bayes for Reconstruction 
data_reconstructed <- list(N=length(I), P=as.integer(n), I=as.integer(I), Z=as.integer(Z), time_added=time_added)

bayes_reconstructed <- stan(model_code = Stan_reconstructed, data=data_reconstructed, iter=500000, chains=4)

#Run Bayes for Simulation
data_sim <- list(N=length(I), P=as.integer(n), I=as.integer(I), S=as.integer(S_sim), time_added=time_added)

bayes_sim <- stan(model_code = Stan_sim, data=data_sim, iter=500000, chains=4)
```
```{r, echo=FALSE}
#Plot S_reconstructed
Sbar <- extract(bayes_reconstructed, 'Sbar')
Sbar <- unlist(Sbar, use.names=FALSE)
lowlim_Sbar <- quantile(Sbar, probs=.025, names=FALSE)
uplim_Sbar <- quantile(Sbar, probs=.975, names=FALSE)

S_lower <- lowlim_Sbar + Z
S_upper <- uplim_Sbar + Z

ggplot(simdata_snoise_mid, aes(x=T))+
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=S_lower, col="S_lower"))+
  geom_line(aes(y=S_upper, col="S_upper"))+
  labs(x="time", y="Number of Susceptibles", title="Reconstructed vs. Simulated Susceptible Dynamics")
```
```{r}
bayes_sim
bayes_reconstructed
```

Adding Noise to the Infected Class:Low Noise
```{r, echo=FALSE}
#Simulate Epidemic Dynamics 
Sim_inoise <- function(beta, n.now, i.now, s.now, r.now, b.now, i.noise) {

results <- as.data.frame(matrix(c(i.now, s.now, r.now, b.now, n.now), nrow=1))
names(results) <- c("I", "S", "R", "B", "N")
  while(s.now*i.now >0){
    i.next <- rnbinom(n=1, mu=(beta*i.now*s.now/n.now)*i.noise, size=i.now) 
    s.now <- s.now - i.next + b.now
    r.now <- r.now + i.now
    b.now <- rgamma(1, shape=2, scale=.000215)*n.now
    n.now <- n.now + b.now 
    i.now <- i.next
    results <- rbind(results, c(i.now, s.now, r.now, b.now, n.now))
  }
return(results)
}

simdata_inoise_low <- Sim_inoise(beta=beta, n.now=n.now, i.now=i.now, s.now=s.now, r.now=r.now, b.now=b.now, i.noise=i.noise.low)

#Pull out Simulation Data and Create Time Vector 
I <- simdata_inoise_low$I
B <- simdata_inoise_low$B
S_sim <- simdata_inoise_low$S
n <- simdata_inoise_low$N
R <- simdata_inoise_low$R
T<- cumsum(rexp(length(I), rate=(1/2.2)))
time_added<-0
for(i in 2:(length(T))){
  time_added[i] <- T[i] - T[i-1]
}


#Plot Simulation
ggplot(simdata_inoise_low, aes(x=T)) +
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=I, col="I"))+
  geom_line(aes(y=R, col="R"))+
  labs(x="Time Steps", y = "Number of Individuals", title = "Simulated Dynamics with Noise in Infected Class, where noise~N(1,.01)")

#Reconstructing Z dynamics 
cumbirths <- cumsum(B)
cuminfecteds <- cumsum(I)
cumulativeinfo <- as.data.frame(matrix(c(cumbirths, cuminfecteds), nrow=length(cuminfecteds)))
names(cumulativeinfo) <- c("cumbirths", "cuminfecteds")

regression <- lm(cumbirths ~ 1, offset=(cuminfecteds))

Z <- as.vector(resid(regression))
```
```{r, include=FALSE}
#Run Bayes
data_reconstructed <- list(N=length(I), P=as.integer(n), I=as.integer(I), Z=as.integer(Z), time_added=time_added)

bayes_reconstructed <- stan(model_code = Stan_reconstructed, data=data_reconstructed, iter=500000, chains=4)

#Run Bayes for Simulation
data_sim <- list(N=length(I), P=as.integer(n), I=as.integer(I), S=as.integer(S_sim), time_added=time_added)

bayes_sim <- stan(model_code = Stan_sim, data=data_sim, iter=500000, chains=4)
```
```{r, echo=FALSE}
#Plot S_reconstructed
Sbar <- extract(bayes_reconstructed, 'Sbar')
Sbar <- unlist(Sbar, use.names=FALSE)
lowlim_Sbar <- quantile(Sbar, probs=.025, names=FALSE)
uplim_Sbar <- quantile(Sbar, probs=.975, names=FALSE)

S_lower <- lowlim_Sbar + Z
S_upper <- uplim_Sbar + Z

ggplot(simdata_inoise_low, aes(x=T))+
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=S_lower, col="S_lower"))+
  geom_line(aes(y=S_upper, col="S_upper"))+
  labs(x="time", y="Number of Susceptibles", title="Reconstructed vs. Simulated Susceptible Dynamics")
```
```{r}
bayes_sim
bayes_reconstructed
```

Adding Noise to the Infected Class:High Noise
```{r, echo=FALSE}
simdata_inoise_high <- Sim_inoise(beta=beta, n.now=n.now, i.now=i.now, s.now=s.now, r.now=r.now, b.now=b.now, i.noise=i.noise.high)

#Pull out Simulation Data and Create Time Vector 
I <- simdata_inoise_high$I
B <- simdata_inoise_high$B
S_sim <- simdata_inoise_high$S
n <- simdata_inoise_high$N
T<- cumsum(rexp(length(I), rate=(1/2.2)))
time_added<-0
for(i in 2:(length(T))){
  time_added[i] <- T[i] - T[i-1]
}

#Plot Simulation
ggplot(simdata_inoise_high, aes(x=T)) +
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=I, col="I"))+
  geom_line(aes(y=R, col="R"))+
  labs(x="Time Steps", y = "Number of Individuals", title = "Simulated Dynamics with Noise in Infected Class, where noise~N(1,.05)")

#Reconstructing Z dynamics 
cumbirths <- cumsum(B)
cuminfecteds <- cumsum(I)
cumulativeinfo <- as.data.frame(matrix(c(cumbirths, cuminfecteds), nrow=length(cuminfecteds)))
names(cumulativeinfo) <- c("cumbirths", "cuminfecteds")

regression <- lm(cumbirths ~ 1, offset=(cuminfecteds))

Z <- as.vector(resid(regression))
```
```{r, include=FALSE}
#Run Bayes
data_reconstructed <- list(N=length(I), P=as.integer(n), I=as.integer(I), Z=as.integer(Z), time_added=time_added)

bayes_reconstructed <- stan(model_code = Stan_reconstructed, data=data_reconstructed, iter=500000, chains=4)

#Run Bayes for Simulation
data_sim <- list(N=length(I), P=as.integer(n), I=as.integer(I), S=as.integer(S_sim), time_added=time_added)

bayes_sim <- stan(model_code = Stan_sim, data=data_sim, iter=500000, chains=4)
```
```{r, echo=FALSE}
#Plot S_reconstructed
Sbar <- extract(bayes_reconstructed, 'Sbar')
Sbar <- unlist(Sbar, use.names=FALSE)
lowlim_Sbar <- quantile(Sbar, probs=.025, names=FALSE)
uplim_Sbar <- quantile(Sbar, probs=.975, names=FALSE)

S_lower <- lowlim_Sbar + Z
S_upper <- uplim_Sbar + Z

ggplot(simdata_inoise_high, aes(x=T))+
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=S_lower, col="S_lower"))+
  geom_line(aes(y=S_upper, col="S_upper"))+
  labs(x="time", y="Number of Susceptibles", title="Reconstructed vs. Simulated Susceptible Dynamics")
```
```{r}
bayes_sim
bayes_reconstructed
```

Adding Noise to the Infected Class:Mid Noise
```{r, echo=FALSE}
simdata_inoise_mid <- Sim_inoise(beta=beta, n.now=n.now, i.now=i.now, s.now=s.now, r.now=r.now, b.now=b.now, i.noise=i.noise.mid)

#Pull out Simulation Data and Create Time Vector 
I <- simdata_inoise_mid$I
B <- simdata_inoise_mid$B
S_sim <- simdata_inoise_mid$S
n <- simdata_inoise_mid$N
T<- cumsum(rexp(length(I), rate=(1/2.2)))
time_added<-0
for(i in 2:(length(T))){
  time_added[i] <- T[i] - T[i-1]
}

#Plot Simulation
ggplot(simdata_inoise_mid, aes(x=T)) +
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=I, col="I"))+
  geom_line(aes(y=R, col="R"))+
  labs(x="Time Steps", y = "Number of Individuals", title = "Simulated Dynamics with Noise in Infected Class, where noise~N(1,.05)")

#Reconstructing Z dynamics 
cumbirths <- cumsum(B)
cuminfecteds <- cumsum(I)
cumulativeinfo <- as.data.frame(matrix(c(cumbirths, cuminfecteds), nrow=length(cuminfecteds)))
names(cumulativeinfo) <- c("cumbirths", "cuminfecteds")

regression <- lm(cumbirths ~ 1, offset=(cuminfecteds))

Z <- as.vector(resid(regression))
```
```{r, include=FALSE}
#Run Bayes
data_reconstructed <- list(N=length(I), P=as.integer(n), I=as.integer(I), Z=as.integer(Z), time_added=time_added)

bayes_reconstructed <- stan(model_code = Stan_reconstructed, data=data_reconstructed, iter=500000, chains=4)

#Run Bayes for Simulation
data_sim <- list(N=length(I), P=as.integer(n), I=as.integer(I), S=as.integer(S_sim), time_added=time_added)

bayes_sim <- stan(model_code = Stan_sim, data=data_sim, iter=500000, chains=4)
```
```{r, echo=FALSE}
#Plot S_reconstructed
Sbar <- extract(bayes_reconstructed, 'Sbar')
Sbar <- unlist(Sbar, use.names=FALSE)
lowlim_Sbar <- quantile(Sbar, probs=.025, names=FALSE)
uplim_Sbar <- quantile(Sbar, probs=.975, names=FALSE)

S_lower <- lowlim_Sbar + Z
S_upper <- uplim_Sbar + Z

ggplot(simdata_inoise_mid, aes(x=T))+
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=S_lower, col="S_lower"))+
  geom_line(aes(y=S_upper, col="S_upper"))+
  labs(x="time", y="Number of Susceptibles", title="Reconstructed vs. Simulated Susceptible Dynamics")
```
```{r}
bayes_sim
bayes_reconstructed
```

Adding Noise to Both Classes
```{r, echo=FALSE}
#Simulate Epidemic Dynamics 
Sim_noise_full <- function(beta, n.now, i.now, s.now, r.now, b.now, i.noise, s.noise) {

results <- as.data.frame(matrix(c(i.now, s.now, r.now, b.now, n.now), nrow=1))
names(results) <- c("I", "S", "R", "B", "N")
  while(s.now*i.now >0){
    i.next <- rnbinom(n=1, mu=(beta*i.now*s.now/n.now)*i.noise, size=i.now) 
    s.now <- s.now - i.next + b.now + s.noise
    r.now <- r.now + i.now
    b.now <- rgamma(1, shape=2, scale=.000215)*n.now
    n.now <- n.now + b.now 
    i.now <- i.next
    results <- rbind(results, c(i.now, s.now, r.now, b.now, n.now))
  }
return(results)
}

simdata_noise_full <- Sim_noise_full(beta=beta, n.now=n.now, i.now=i.now, s.now=s.now, r.now=r.now, b.now=b.now, i.noise=i.noise.low, s.noise=s.noise.low)

#Pull out Simulation Data and Create Time Vector 
I <- simdata_noise_full$I
B <- simdata_noise_full$B
S_sim <- simdata_noise_full$S
n <- simdata_noise_full$N
T<- cumsum(rexp(length(I), rate=(1/2.2)))
time_added<-0
for(i in 2:(length(T))){
  time_added[i] <- T[i] - T[i-1]
}

#Plot Simulation
ggplot(simdata_noise_full, aes(x=T)) +
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=I, col="I"))+
  geom_line(aes(y=R, col="R"))+
  labs(x="Time Steps", y = "Number of Individuals", title = "Simulated Dynamics with Low Noise in Both Classes")

#Reconstructing Z dynamics 
cumbirths <- cumsum(B)
cuminfecteds <- cumsum(I)
cumulativeinfo <- as.data.frame(matrix(c(cumbirths, cuminfecteds), nrow=length(cuminfecteds)))
names(cumulativeinfo) <- c("cumbirths", "cuminfecteds")

regression <- lm(cumbirths ~ 1, offset=(cuminfecteds))

Z <- as.vector(resid(regression))
```
```{r, include=FALSE}
#Run Bayes
data_reconstructed <- list(N=length(I), P=as.integer(n), I=as.integer(I), Z=as.integer(Z), time_added=time_added)

bayes_reconstructed <- stan(model_code = Stan_reconstructed, data=data_reconstructed, iter=500000, chains=4)

#Run Bayes for Simulation
data_sim <- list(N=length(I), P=as.integer(n), I=as.integer(I), S=as.integer(S_sim), time_added=time_added)

bayes_sim <- stan(model_code = Stan_sim, data=data_sim, iter=500000, chains=4)
```
```{r, echo=FALSE}
#Plot S_reconstructed
Sbar <- extract(bayes_reconstructed, 'Sbar')
Sbar <- unlist(Sbar, use.names=FALSE)
lowlim_Sbar <- quantile(Sbar, probs=.025, names=FALSE)
uplim_Sbar <- quantile(Sbar, probs=.975, names=FALSE)

S_lower <- lowlim_Sbar + Z
S_upper <- uplim_Sbar + Z

ggplot(simdata_noise_full, aes(x=T))+
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=S_lower, col="S_lower"))+
  geom_line(aes(y=S_upper, col="S_upper"))+
  labs(x="time", y="Number of Susceptibles", title="Reconstructed vs. Simulated Susceptible Dynamics")
```
```{r}
bayes_sim
bayes_reconstructed
```



