---
title: "Notes"
author: "Kaitlyn Stocker"
date: "4/6/2017"
output: html_document
---

```{r Meeting Notes 3/30/17}
#meeting notes 3/30/17
# I have their betas, I know what model - don't know what rho's they fit 
# can I reverse engineer to figure out rho's?? 
#one page summary I can write up on local regression to get rho's (Dr haran can ask Matt Ferrari) -- in particular, roughly speaking 1 pg, methodology used in order to get the rho's I get. Add some lil graphs. SEND WITHIN A WEEK 

#first run analysis through all of the cities
#then talk about what's happening - what's similar, what's different... 
#what's interesting across these cities? 
#how to model stuff across cities? (really clever -not spatial model.. things that might be happening synchronicely or a-synchronicely)

#Pick 3 places and try to learn stuff across those 3 places (northeast?) - what have they written about these places - also look at places that are smaller than the big guys (at least 2 of the smaller guys). Compare across... how well things fit for the certain city... ?Proportion of 0 inferred infecteds?...are there die-outs? 

#Look at paper: jandarov (on dr haran's website) on the gravity TSIR model. Get a taste of spatial version. Cool thing: 4 signitures (signiture=characteristic of how an infectious disease plays otu) - one is the propensity to get snuffed out. THey wrote a model that reproduced those characteristics well. *does my fitted model have the same signitures as the model itself 

#forward spatial model is possible - but the inference might be too far. May even be able to do the gravity model 
```

To do (4/6/17):
- write one page summary of my SR with seasonal rho's 
- keep reading Jandarov and Haran paper on TSIR gravity model 
- troubleshoot bayes issues (phil, boston, stlouis, etc)
  - anything common between them? (check: paper betas, population, SR and IR graphs)
- add names of graphs to graphing functions 
- add new code to deal with NA's 
  - see if ignoring NA's in cumsum functions fixes the problem, or if greater problems persist     (make sure to remove from both births and cases for same datapoint)
  - are NA's die outs???
- Does switching the cumbirths and cumcases in the regression make a difference? maybe try to regress the other way 


```{r}

```

Questions for Dr Haran 4/6/17:

- what is the best way to deal with NA's? I don't want to delete them because it will shift my biweeks, and there is no "na.rm" or other option in the cumsum function, which is necessary to reconstruction. For cities with strings of only 1 NA in a row, I took the average of the value before and after it to fill it in - is this an okay thing to do? Feels fudgy but I don't feel like it should impact analysis 
  -answer: all of the missing values are now new parameters to estimate in the model (in bayes)- need to make sure that there's a way for the model to learn about the missing values from the data
  
  
- some cities run fine through everything, and some come up with errors for bayes. I haven't been able to troubleshoot what the matter is yet. 

- The betas in your Jandarov paper are much different from the betas in the paper I am working from. Your betas are around 1, the betas that I am working with are in the 20s through 50s. Both are for measles data - why would this be? 

```{r}
#meeting notes 4/6/17
#make a list of strange anomalies (betas, etc)

#Beta mystery: are there variations in betas for the same data sets? What is my Rnot? 

#figure out if NA's are actually zeros
#find distribution of preceding values, distribution of post values, is it decreasing and then increasing? etc. If they are zeros, input 0.5 or 1 or something similar to start 
#use interpolator or -- build an empirical conditional distribution so look at what's happening around an NA, find places in data where things look similar but withouth NA's and assume that the same thing would be about the same 

#not from dr haran but - what if I used NA's to indicate when a spark comes to re-ignite the infection? could maybe use that to find some kind of baby spatial stuff 
```


4/27/17:
-Talk about parameterizing beta to match with timestep (noteable when switching from deterministic to chain binomial: switch from daily timestep to timestep being equal to the infectious period. Therefore must multiply beta by the length of the infectious period in days)
- fix figure captions, figure references
- add citiations
- move bayesian and MLE inference to their own section


4/27 Meeting Notes:
- Next week: ben will be gone 
- NO MEETING 18th and 25th (resume June 1)
- if I find Ottar: ask him about reporting rates
- could put together code and examples and put it on github (vignette of infetious disease)


THESIS WRITING TO DO:
- Section on time varying rho/local regression (started in latex)
- Section about time varying reporting rates, oscillatory dynamics (started in latex)
- In deterministic and/or chain binomial section, add paragraph talking about parameterizing beta to match with timestep 
- continue adding plots, re-work examples to be consistent, add citations 

```{r}
#Thesis Detail Work:
- fix axis labels on the MLE plot for chain binomial (and check other axis labels)
- fix notation in seasonal rho section 
- fix figure sizing and captioning 
- fix figure references
- add more citations 
```

Questions for Dr. Haran:
-notation for seasonal rho section
-what might Finkenstadt and Grenfell mean when they say they chose a bandwidth of .3 times the standard deviation of Xt when sd(cumcases) = 128389.4?s


To do (4/6/17):
- write one page summary of my SR with seasonal rho's 
- keep reading Jandarov and Haran paper on TSIR gravity model 
- troubleshoot bayes issues (phil, boston, stlouis, etc)
  - anything common between them? (check: paper betas, population, SR and IR graphs)
  - priors on beta and sbar - can they be more general? try to find problem 
- add names of graphs to graphing functions 
- add new code to deal with NA's 
  - see if ignoring NA's in cumsum functions fixes the problem, or if greater problems persist     (make sure to remove from both births and cases for same datapoint)
- Does switching the cumbirths and cumcases in the regression make a difference? maybe try to regress the other way 





Questions for Dr Haran 4/6/17:

- what is the best way to deal with NA's? I don't want to delete them because it will shift my biweeks, and there is no "na.rm" or other option in the cumsum function, which is necessary to reconstruction. For cities with strings of only 1 NA in a row, I took the average of the value before and after it to fill it in - is this an okay thing to do? Feels fudgy but I don't feel like it should impact analysis 

  - if I remove NA's or ignore NA cases when computing cumulative sums, should I also exclude the birth data at those times so that cumulative births and cumulative cases are in line with one another? 
  
- some cities run fine through everything, and some come up with errors for bayes. I haven't been able to troubleshoot what the matter is yet. 




Meeting 5/11:
- For forward simulation: use the entire markov chain sample for my betas (maybe every thousandth). FOr each subsampled makrov chain, run simulation, at each time point take the median beta (for 1000+ curves)
- look at posterior means and standard errors for betas from bayes ~ use mcse package in R (monte carlo standard error). Use batchmeans, for computing standard errors. Take matrix of betas and make sure that the standard errors of posterior mean estimates are small (Stan probs does this for me, double check)
- is there something inherently different between how measles behaves in big cities vs small cities vs medium cities ~ start saying things qualitatively, then start defining what is really happening quantitatively 

#Questions for June 1st Meeting:
- when sampling the markov chain, not sure what you meant by taking the median beta at each time point for the 1000+ curves. Was able to sample, simulate, and plot 1000 curves, but wouldn't the median beta for each time point just be the same thing I already did? Maybe I misunderstood.
- Will be away June 15th (2 weeks from this weeks meeting)
- how to make plot like the one from the paper (for susceptible vs infected proportions)?
  - note: they chose initial conditions based on the distance between the simulation and the data during the peak month of may. I did it for all datapoints. 


#June 1st meeting
- with 1000 curves, can get 95% interval for the number of infecteds at each time point. Make plot similar to my old susceptible reconstruction plots (learning about hierarchical distributions and mixture modeling)... similar to my plot but take 95% interval for each 
- kernel density 2d plot (go on cran and look it up)
-free time series book: Cryer and Chan called Time Series Analysis with Applications in R (on Dr Haran's website, go to courses taught in previous semesters)
- reason Bjornstad et al chose to do MSE for just peak in may vs whole curve is to capture disease signitures ~ could try that myself 
- what next?: 
  - get thesis DONE by July 1st. 
  - short qualitative essay on what is the difference between small/medium/large cities ~ how does city size affect model fit? why? what's happening? 
  ***EMAIL DR HARAN the outline we talked about***
  - tuesday june 13th 1:30-3:30 (optimistic)intensive thesis writeup meeting
  - chapter intros: extremely short paragraphs (5 sentences) that says what the chapter is about (maybe make a sketch of the intros first, should tell someone what I did)
  
#Notes for June 8 meeting 
- when I took 95% interval for each time point, some of them did not include the original simulation with the median betas. 
  - options: 
      - overwrite interval where it doesn't include the og simulation to have the           bound be at the simulation.
      - change method of computing intervals to be a confidence interval with the og         simulation as the mean. 
      - ??
-