---
title: "Notes"
author: "Kaitlyn Stocker"
date: "4/6/2017"
output: html_document
---

```{r Meeting Notes 3/30/17}
#meeting notes 3/30/17
# I have their betas, I know what model - don't know what rho's they fit 
# can I reverse engineer to figure out rho's?? 
#one page summary I can write up on local regression to get rho's (Dr haran can ask Matt Ferrari) -- in particular, roughly speaking 1 pg, methodology used in order to get the rho's I get. Add some lil graphs. SEND WITHIN A WEEK 

#first run analysis through all of the cities
#then talk about what's happening - what's similar, what's different... 
#what's interesting across these cities? 
#how to model stuff across cities? (really clever -not spatial model.. things that might be happening synchronicely or a-synchronicely)

#Pick 3 places and try to learn stuff across those 3 places (northeast?) - what have they written about these places - also look at places that are smaller than the big guys (at least 2 of the smaller guys). Compare across... how well things fit for the certain city... ?Proportion of 0 inferred infecteds?...are there die-outs? 

#Look at paper: jandarov (on dr haran's website) on the gravity TSIR model. Get a taste of spatial version. Cool thing: 4 signitures (signiture=characteristic of how an infectious disease plays otu) - one is the propensity to get snuffed out. THey wrote a model that reproduced those characteristics well. *does my fitted model have the same signitures as the model itself 

#forward spatial model is possible - but the inference might be too far. May even be able to do the gravity model 
```

To do (4/6/17):
- write one page summary of my SR with seasonal rho's 
- keep reading Jandarov and Haran paper on TSIR gravity model 
- troubleshoot bayes issues (phil, boston, stlouis, etc)
  - anything common between them? (check: paper betas, population, SR and IR graphs)
- add names of graphs to graphing functions 
- add new code to deal with NA's 
  - see if ignoring NA's in cumsum functions fixes the problem, or if greater problems persist     (make sure to remove from both births and cases for same datapoint)
  - are NA's die outs???
- Does switching the cumbirths and cumcases in the regression make a difference? maybe try to regress the other way 


```{r}

```

Questions for Dr Haran 4/6/17:

- what is the best way to deal with NA's? I don't want to delete them because it will shift my biweeks, and there is no "na.rm" or other option in the cumsum function, which is necessary to reconstruction. For cities with strings of only 1 NA in a row, I took the average of the value before and after it to fill it in - is this an okay thing to do? Feels fudgy but I don't feel like it should impact analysis 
  -answer: all of the missing values are now new parameters to estimate in the model (in bayes)- need to make sure that there's a way for the model to learn about the missing values from the data
  
  
- some cities run fine through everything, and some come up with errors for bayes. I haven't been able to troubleshoot what the matter is yet. 

- The betas in your Jandarov paper are much different from the betas in the paper I am working from. Your betas are around 1, the betas that I am working with are in the 20s through 50s. Both are for measles data - why would this be? 

```{r}
#meeting notes 4/6/17
#make a list of strange anomalies (betas, etc)

#Beta mystery: are there variations in betas for the same data sets? What is my Rnot? 

#figure out if NA's are actually zeros
#find distribution of preceding values, distribution of post values, is it decreasing and then increasing? etc. If they are zeros, input 0.5 or 1 or something similar to start 
#use interpolator or -- build an empirical conditional distribution so look at what's happening around an NA, find places in data where things look similar but withouth NA's and assume that the same thing would be about the same 

#not from dr haran but - what if I used NA's to indicate when a spark comes to re-ignite the infection? could maybe use that to find some kind of baby spatial stuff 
```


4/27/17:
-Talk about parameterizing beta to match with timestep (noteable when switching from deterministic to chain binomial: switch from daily timestep to timestep being equal to the infectious period. Therefore must multiply beta by the length of the infectious period in days)
- fix figure captions, figure references
- add citiations
- move bayesian and MLE inference to their own section


4/27 Meeting Notes:
- Next week: ben will be gone 
- NO MEETING 18th and 25th (resume June 1)
- if I find Ottar: ask him about reporting rates
- could put together code and examples and put it on github (vignette of infetious disease)


THESIS WRITING TO DO:
- Section on time varying rho/local regression (started in latex)
- Section about time varying reporting rates, oscillatory dynamics (started in latex)
- In deterministic and/or chain binomial section, add paragraph talking about parameterizing beta to match with timestep 
- continue adding plots, re-work examples to be consistent, add citations 
