---
title: "Introducing Noise"
author: "Kaitlyn Stocker"
date: "2/6/2017"
output: html_document
---

```{r, include=FALSE}
library(rstan)
library(ggplot2)
```

Small Sample Size (N_0 = 76,300)
```{r, echo=FALSE}
#Simulation Function of TSIR Model 
Sim <- function(beta, n.now, i.now, s.now, r.now, b.now) {

results <- as.data.frame(matrix(c(i.now, s.now, r.now, b.now, n.now), nrow=1))
names(results) <- c("I", "S", "R", "B", "N")
  while(s.now*i.now >0){
    i.next <- rnbinom(n=1, mu=beta*i.now*s.now/n.now, size=i.now) 
    s.now <- s.now - i.next + b.now
    r.now <- r.now + i.now
    b.now <- rgamma(1, shape=2, scale=.000215)*n.now
    n.now <- n.now + b.now 
    i.now <- i.next
    results <- rbind(results, c(i.now, s.now, r.now, b.now, n.now))
  }
return(results)
}

#Initial Conditions 
n.now <- 763000
s.now <- 762995
i.now <- n.now-s.now
r.now <- 0
beta <- 1.66
birthrate <- rgamma(1, shape=2, scale=.000215)
b.now <- 0


#Run Simulation
simdata <- Sim(beta=beta, n.now=n.now, i.now=i.now, s.now=s.now, r.now=r.now, b.now=b.now)

#Define Time Steps
T<- cumsum(rexp(length(simdata$I), rate=(1/2.2)))

#Plot Simulation
ggplot(simdata, aes(x=T)) +
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=I, col="I"))+
  geom_line(aes(y=R, col="R"))+
  labs(x="Time Steps", y = "Number of Individuals", title = "Simulated Epidemic Dynamics")



```
```{r, echo=FALSE}
#Reconstructing Z dynamics 
cumbirths <- cumsum(simdata$B)
cuminfecteds <- cumsum(simdata$I)
cumulativeinfo <- as.data.frame(matrix(c(cumbirths, cuminfecteds), nrow=length(cuminfecteds)))
names(cumulativeinfo) <- c("cumbirths", "cuminfecteds")


#plot(cumbirths ~ cuminfecteds)

regression <- lm(cumbirths ~ 1, offset=(cuminfecteds))
#summary(regression)

Z <- as.vector(resid(regression))
#summary(Z)

ggplot(simdata, aes(x=T))+
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=Z, col="Z"))+
  labs(x="Time", y="Number of Susceptibles", title="Reconstructed Dynamics")
```
```{r, include=FALSE}
#Preparing Data for Bayes
I<- simdata$I
n <- simdata$N

time_added<-0
for(i in 2:(length(T))){
  time_added[i] <- T[i] - T[i-1]
}

data_z <- list(N=length(I), P=as.integer(n), I=as.integer(I), Z=as.integer(Z), time_added=time_added)

```
```{r, include=FALSE}
Stan_z <- as.character("
data {
	int<lower=1> N; // time steps
  real P[N]; //population size
  real time_added[N];
	int <lower=0> I[N]; // infecteds 
	int  Z[N]; // Zt 
}

parameters {
  real <lower=P[1]/100, upper= P[N]> Sbar;
	real<lower=0, upper=5> beta;
  real<lower=0, upper=5> gamma;
}

model {
	//priors
  Sbar ~ normal(P[1]/2, P[1]/10);
	beta ~ gamma(3, 1);
  gamma ~ gamma(2,1);

	//likelihood
	for (n in 2:N){
		I[n] ~ neg_binomial_2(beta*I[n-1]*(Z[n-1]+Sbar)/P[n-1], I[n-1]);
    time_added[n]~exponential_log(gamma);
  }
}

                            ")

bayes_reconstructed <- stan(model_code = Stan_z, data=data_z, iter=500000, chains=4)
```
```{r, echo=FALSE}
Sbar <- extract(bayes_reconstructed, 'Sbar')
Sbar <- unlist(Sbar, use.names=FALSE)
lowlim_Sbar <- quantile(Sbar, probs=.025, names=FALSE)
uplim_Sbar <- quantile(Sbar, probs=.975, names=FALSE)

S_lower <- lowlim_Sbar + Z
S_upper <- uplim_Sbar + Z

ggplot(simdata, aes(x=T))+
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=S_lower, col="S_lower"))+
  geom_line(aes(y=S_upper, col="S_upper"))+
  labs(x="time", y="Number of Susceptibles", title="Reconstructed vs. Simulated Susceptible Dynamics")
```
```{r, include=FALSE}
##Running Bayes with True Susceptible Dynamics 
S_sim <- simdata$S

data_sim <- list(N=length(I), P=as.integer(n), I=as.integer(I), S=as.integer(S_sim), time_added=time_added)

SIRstancode <- as.character("
                            data {
	int<lower=1> N; // time steps
  real P[N]; //population size
  real time_added[N];
	int <lower=0> I[N];
	int <lower=0> S[N];
}


parameters {
	real<lower=0, upper=5> beta;
  real<lower=0, upper=5> gamma;

}

model {
	//priors
	beta ~ gamma(3, 1);
  gamma ~ gamma(2,1);
	//likelihood
	for (n in 2:N){
		I[n] ~ neg_binomial_2(beta*I[n-1]*S[n-1]/P[n-1], I[n-1]);
    time_added[n]~exponential_log(gamma);
  }
}

                            ")

bayes_sim <- stan(model_code = SIRstancode, data=data_sim, iter=500000, chains=4)
```
```{r}
bayes_sim
bayes_reconstructed
```



Adding Randomness to Susceptible Class 
```{r, echo=FALSE}
#Simulation Function of TSIR Model 
Sim <- function(beta, n.now, i.now, s.now, r.now, b.now) {

results <- as.data.frame(matrix(c(i.now, s.now, r.now, b.now, n.now), nrow=1))
names(results) <- c("I", "S", "R", "B", "N")
  while(s.now*i.now >0){
    i.next <- rnbinom(n=1, mu=beta*i.now*s.now/n.now, size=i.now) 
    s.now <- s.now - i.next + b.now + rnorm(1,0,n.now/500)
    r.now <- r.now + i.now
    b.now <- rgamma(1, shape=2, scale=.000215)*n.now
    n.now <- n.now + b.now 
    i.now <- i.next
    results <- rbind(results, c(i.now, s.now, r.now, b.now, n.now))
  }
return(results)
}

#Initial Conditions 
n.now <- 763000
s.now <- 762995
i.now <- n.now-s.now
r.now <- 0
beta <- 1.66
birthrate <- rgamma(1, shape=2, scale=.000215)
b.now <- 0


#Run Simulation
simdata <- Sim(beta=beta, n.now=n.now, i.now=i.now, s.now=s.now, r.now=r.now, b.now=b.now)

#Define Time Steps
T<- cumsum(rexp(length(simdata$I), rate=(1/2.2)))

#Plot Simulation
ggplot(simdata, aes(x=T)) +
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=I, col="I"))+
  geom_line(aes(y=R, col="R"))+
  labs(x="Time Steps", y = "Number of Individuals", title = "Simulated Epidemic Dynamics")



```
```{r, echo=FALSE}
#Reconstructing Z dynamics 
cumbirths <- cumsum(simdata$B)
cuminfecteds <- cumsum(simdata$I)
cumulativeinfo <- as.data.frame(matrix(c(cumbirths, cuminfecteds), nrow=length(cuminfecteds)))
names(cumulativeinfo) <- c("cumbirths", "cuminfecteds")


#plot(cumbirths ~ cuminfecteds)

regression <- lm(cumbirths ~ 1, offset=(cuminfecteds))
#summary(regression)

Z <- as.vector(resid(regression))
#summary(Z)

ggplot(simdata, aes(x=T))+
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=Z, col="Z"))+
  labs(x="Time", y="Number of Susceptibles", title="Reconstructed Dynamics")
```
```{r, include=FALSE}
#Preparing Data for Bayes
I<- simdata$I
n <- simdata$N

time_added<-0
for(i in 2:(length(T))){
  time_added[i] <- T[i] - T[i-1]
}

data_z <- list(N=length(I), P=as.integer(n), I=as.integer(I), Z=as.integer(Z), time_added=time_added)

```
```{r, include=FALSE}
Stan_z <- as.character("
data {
	int<lower=1> N; // time steps
  real P[N]; //population size
  real time_added[N];
	int <lower=0> I[N]; // infecteds 
	int  Z[N]; // Zt 
}

parameters {
  real <lower=P[1]/100, upper= P[N]> Sbar;
	real<lower=0, upper=5> beta;
  real<lower=0, upper=5> gamma;
}

model {
	//priors
  Sbar ~ normal(P[1]/2, P[1]/10);
	beta ~ gamma(3, 1);
  gamma ~ gamma(2,1);

	//likelihood
	for (n in 2:N){
		I[n] ~ neg_binomial_2(beta*I[n-1]*(Z[n-1]+Sbar)/P[n-1], I[n-1]);
    time_added[n]~exponential_log(gamma);
  }
}

                            ")

bayes_reconstructed <- stan(model_code = Stan_z, data=data_z, iter=500000, chains=4)
```
```{r, echo=FALSE}
Sbar <- extract(bayes_reconstructed, 'Sbar')
Sbar <- unlist(Sbar, use.names=FALSE)
lowlim_Sbar <- quantile(Sbar, probs=.025, names=FALSE)
uplim_Sbar <- quantile(Sbar, probs=.975, names=FALSE)

S_lower <- lowlim_Sbar + Z
S_upper <- uplim_Sbar + Z

ggplot(simdata, aes(x=T))+
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=S_lower, col="S_lower"))+
  geom_line(aes(y=S_upper, col="S_upper"))+
  labs(x="time", y="Number of Susceptibles", title="Reconstructed vs. Simulated Susceptible Dynamics")
```
```{r, include=FALSE}
##Running Bayes with True Susceptible Dynamics 
S_sim <- simdata$S

data_sim <- list(N=length(I), P=as.integer(n), I=as.integer(I), S=as.integer(S_sim), time_added=time_added)

SIRstancode <- as.character("
                            data {
	int<lower=1> N; // time steps
  real P[N]; //population size
  real time_added[N];
	int <lower=0> I[N];
	int <lower=0> S[N];
}


parameters {
	real<lower=0, upper=5> beta;
  real<lower=0, upper=5> gamma;

}

model {
	//priors
	beta ~ gamma(3, 1);
  gamma ~ gamma(2,1);
	//likelihood
	for (n in 2:N){
		I[n] ~ neg_binomial_2(beta*I[n-1]*S[n-1]/P[n-1], I[n-1]);
    time_added[n]~exponential_log(gamma);
  }
}

                            ")

bayes_sim <- stan(model_code = SIRstancode, data=data_sim, iter=500000, chains=4)
```
```{r}
bayes_sim
bayes_reconstructed
```



Adding Noise to Infected Class
```{r, echo=FALSE}
#Simulation Function of TSIR Model 
Sim <- function(beta, n.now, i.now, s.now, r.now, b.now) {

results <- as.data.frame(matrix(c(i.now, s.now, r.now, b.now, n.now), nrow=1))
names(results) <- c("I", "S", "R", "B", "N")
  while(s.now*i.now >0){
    i.next <- rnbinom(n=1, mu=(beta*i.now*s.now/n.now)*abs(rnorm(1,1,.01)), size=i.now) 
    s.now <- s.now - i.next + b.now
    r.now <- r.now + i.now
    b.now <- rgamma(1, shape=2, scale=.000215)*n.now
    n.now <- n.now + b.now 
    i.now <- i.next
    results <- rbind(results, c(i.now, s.now, r.now, b.now, n.now))
  }
return(results)
}

#Initial Conditions 
n.now <- 763000
s.now <- 762995
i.now <- n.now-s.now
r.now <- 0
beta <- 1.66
birthrate <- rgamma(1, shape=2, scale=.000215)
b.now <- 0


#Run Simulation
simdata <- Sim(beta=beta, n.now=n.now, i.now=i.now, s.now=s.now, r.now=r.now, b.now=b.now)

#Define Time Steps
T<- cumsum(rexp(length(simdata$I), rate=(1/2.2)))

#Plot Simulation
ggplot(simdata, aes(x=T)) +
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=I, col="I"))+
  geom_line(aes(y=R, col="R"))+
  labs(x="Time Steps", y = "Number of Individuals", title = "Simulated Epidemic Dynamics")



```
```{r, echo=FALSE}
#Reconstructing Z dynamics 
cumbirths <- cumsum(simdata$B)
cuminfecteds <- cumsum(simdata$I)
cumulativeinfo <- as.data.frame(matrix(c(cumbirths, cuminfecteds), nrow=length(cuminfecteds)))
names(cumulativeinfo) <- c("cumbirths", "cuminfecteds")


#plot(cumbirths ~ cuminfecteds)

regression <- lm(cumbirths ~ 1, offset=(cuminfecteds))
#summary(regression)

Z <- as.vector(resid(regression))
#summary(Z)

ggplot(simdata, aes(x=T))+
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=Z, col="Z"))+
  labs(x="Time", y="Number of Susceptibles", title="Reconstructed Dynamics")
```
```{r, include=FALSE}
#Preparing Data for Bayes
I<- simdata$I
n <- simdata$N

time_added<-0
for(i in 2:(length(T))){
  time_added[i] <- T[i] - T[i-1]
}

data_z <- list(N=length(I), P=as.integer(n), I=as.integer(I), Z=as.integer(Z), time_added=time_added)

```
```{r, include=FALSE}
Stan_z <- as.character("
data {
	int<lower=1> N; // time steps
  real P[N]; //population size
  real time_added[N];
	int <lower=0> I[N]; // infecteds 
	int  Z[N]; // Zt 
}

parameters {
  real <lower=P[1]/100, upper= P[N]> Sbar;
	real<lower=0, upper=5> beta;
  real<lower=0, upper=5> gamma;
}

model {
	//priors
  Sbar ~ normal(P[1]/2, P[1]/10);
	beta ~ gamma(3, 1);
  gamma ~ gamma(2,1);

	//likelihood
	for (n in 2:N){
		I[n] ~ neg_binomial_2(beta*I[n-1]*(Z[n-1]+Sbar)/P[n-1], I[n-1]);
    time_added[n]~exponential_log(gamma);
  }
}

                            ")

bayes_reconstructed <- stan(model_code = Stan_z, data=data_z, iter=500000, chains=4)
```
```{r, echo=FALSE}
Sbar <- extract(bayes_reconstructed, 'Sbar')
Sbar <- unlist(Sbar, use.names=FALSE)
lowlim_Sbar <- quantile(Sbar, probs=.025, names=FALSE)
uplim_Sbar <- quantile(Sbar, probs=.975, names=FALSE)

S_lower <- lowlim_Sbar + Z
S_upper <- uplim_Sbar + Z

ggplot(simdata, aes(x=T))+
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=S_lower, col="S_lower"))+
  geom_line(aes(y=S_upper, col="S_upper"))+
  labs(x="time", y="Number of Susceptibles", title="Reconstructed vs. Simulated Susceptible Dynamics")
```
```{r, include=FALSE}
##Running Bayes with True Susceptible Dynamics 
S_sim <- simdata$S

data_sim <- list(N=length(I), P=as.integer(n), I=as.integer(I), S=as.integer(S_sim), time_added=time_added)

SIRstancode <- as.character("
                            data {
	int<lower=1> N; // time steps
  real P[N]; //population size
  real time_added[N];
	int <lower=0> I[N];
	int <lower=0> S[N];
}


parameters {
	real<lower=0, upper=5> beta;
  real<lower=0, upper=5> gamma;

}

model {
	//priors
	beta ~ gamma(3, 1);
  gamma ~ gamma(2,1);
	//likelihood
	for (n in 2:N){
		I[n] ~ neg_binomial_2(beta*I[n-1]*S[n-1]/P[n-1], I[n-1]);
    time_added[n]~exponential_log(gamma);
  }
}

                            ")

bayes_sim <- stan(model_code = SIRstancode, data=data_sim, iter=500000, chains=4)
```
```{r}
bayes_sim
bayes_reconstructed
```

Adding Noise to Infecteds with Jitter
```{r, echo=FALSE}
#Simulation Function of TSIR Model 
Sim <- function(beta, n.now, i.now, s.now, r.now, b.now) {

results <- as.data.frame(matrix(c(i.now, s.now, r.now, b.now, n.now), nrow=1))
names(results) <- c("I", "S", "R", "B", "N")
  while(s.now*i.now >0){
    i.next <- rnbinom(n=1, mu=beta*i.now*s.now/n.now, size=i.now) 
    s.now <- s.now - i.next + b.now
    r.now <- r.now + i.now
    b.now <- rgamma(1, shape=2, scale=.000215)*n.now
    n.now <- n.now + b.now 
    i.now <- i.next
    results <- rbind(results, c(i.now, s.now, r.now, b.now, n.now))
  }
return(results)
}

#Initial Conditions 
n.now <- 763000
s.now <- 762995
i.now <- n.now-s.now
r.now <- 0
beta <- 1.66
birthrate <- rgamma(1, shape=2, scale=.000215)
b.now <- 0


#Run Simulation
simdata <- Sim(beta=beta, n.now=n.now, i.now=i.now, s.now=s.now, r.now=r.now, b.now=b.now)

#Define Time Steps
T<- cumsum(rexp(length(simdata$I), rate=(1/2.2)))

#Plot Simulation
ggplot(simdata, aes(x=T)) +
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=I, col="I"))+
  geom_line(aes(y=R, col="R"))+
  labs(x="Time Steps", y = "Number of Individuals", title = "Simulated Epidemic Dynamics")



```
```{r}
#Add Jitter to Infecteds 
I<- simdata$I
I_jitter <- jitter(I)

ggplot(simdata, aes(x=T)) +
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=I_jitter, col="I_jitter"))+
  geom_line(aes(y=R, col="R"))+
  labs(x="Time Steps", y = "Number of Individuals", title = "Simulated Epidemic Dynamics")
```
```{r, echo=FALSE}
#Reconstructing Z dynamics 
cumbirths <- cumsum(simdata$B)
cuminfecteds <- cumsum(I_jitter)
cumulativeinfo <- as.data.frame(matrix(c(cumbirths, cuminfecteds), nrow=length(cuminfecteds)))
names(cumulativeinfo) <- c("cumbirths", "cuminfecteds")


#plot(cumbirths ~ cuminfecteds)

regression <- lm(cumbirths ~ 1, offset=(cuminfecteds))
#summary(regression)

Z <- as.vector(resid(regression))
#summary(Z)

ggplot(simdata, aes(x=T))+
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=Z, col="Z"))+
  labs(x="Time", y="Number of Susceptibles", title="Reconstructed Dynamics")
```
```{r, include=FALSE}
#Preparing Data for Bayes
n <- simdata$N

time_added<-0
for(i in 2:(length(T))){
  time_added[i] <- T[i] - T[i-1]
}

data_z <- list(N=length(I_jitter), P=as.integer(n), I=as.integer(I_jitter), Z=as.integer(Z), time_added=time_added)

```
```{r, include=FALSE}
Stan_z <- as.character("
data {
	int<lower=1> N; // time steps
  real P[N]; //population size
  real time_added[N];
	int <lower=0> I[N]; // infecteds 
	int  Z[N]; // Zt 
}

parameters {
  real <lower=P[1]/100, upper= P[N]> Sbar;
	real<lower=0, upper=5> beta;
  real<lower=0, upper=5> gamma;
}

model {
	//priors
  Sbar ~ normal(P[1]/2, P[1]/10);
	beta ~ gamma(3, 1);
  gamma ~ gamma(2,1);

	//likelihood
	for (n in 2:N){
		I[n] ~ neg_binomial_2(beta*I[n-1]*(Z[n-1]+Sbar)/P[n-1], I[n-1]);
    time_added[n]~exponential_log(gamma);
  }
}

                            ")

bayes_reconstructed <- stan(model_code = Stan_z, data=data_z, iter=500000, chains=4)
```
```{r, echo=FALSE}
Sbar <- extract(bayes_reconstructed, 'Sbar')
Sbar <- unlist(Sbar, use.names=FALSE)
lowlim_Sbar <- quantile(Sbar, probs=.025, names=FALSE)
uplim_Sbar <- quantile(Sbar, probs=.975, names=FALSE)

S_lower <- lowlim_Sbar + Z
S_upper <- uplim_Sbar + Z

ggplot(simdata, aes(x=T))+
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=S_lower, col="S_lower"))+
  geom_line(aes(y=S_upper, col="S_upper"))+
  labs(x="time", y="Number of Susceptibles", title="Reconstructed vs. Simulated Susceptible Dynamics")
```
```{r, include=FALSE}
##Running Bayes with True Susceptible Dynamics 
S_sim <- simdata$S

data_sim <- list(N=length(I), P=as.integer(n), I=as.integer(I), S=as.integer(S_sim), time_added=time_added)

SIRstancode <- as.character("
                            data {
	int<lower=1> N; // time steps
  real P[N]; //population size
  real time_added[N];
	int <lower=0> I[N];
	int <lower=0> S[N];
}


parameters {
	real<lower=0, upper=5> beta;
  real<lower=0, upper=5> gamma;

}

model {
	//priors
	beta ~ gamma(3, 1);
  gamma ~ gamma(2,1);
	//likelihood
	for (n in 2:N){
		I[n] ~ neg_binomial_2(beta*I[n-1]*S[n-1]/P[n-1], I[n-1]);
    time_added[n]~exponential_log(gamma);
  }
}

                            ")

bayes_sim <- stan(model_code = SIRstancode, data=data_sim, iter=500000, chains=4)
```
```{r}
bayes_sim
bayes_reconstructed
```{r}

```


Adding Noise to Susceptibles with Jitter
```{r, echo=FALSE}
#Simulation Function of TSIR Model 
Sim <- function(beta, n.now, i.now, s.now, r.now, b.now) {

results <- as.data.frame(matrix(c(i.now, s.now, r.now, b.now, n.now), nrow=1))
names(results) <- c("I", "S", "R", "B", "N")
  while(s.now*i.now >0){
    i.next <- rnbinom(n=1, mu=beta*i.now*s.now/n.now, size=i.now) 
    s.now <- s.now - i.next + b.now
    r.now <- r.now + i.now
    b.now <- rgamma(1, shape=2, scale=.000215)*n.now
    n.now <- n.now + b.now 
    i.now <- i.next
    results <- rbind(results, c(i.now, s.now, r.now, b.now, n.now))
  }
return(results)
}

#Initial Conditions 
n.now <- 763000
s.now <- 762995
i.now <- n.now-s.now
r.now <- 0
beta <- 1.66
birthrate <- rgamma(1, shape=2, scale=.000215)
b.now <- 0


#Run Simulation
simdata <- Sim(beta=beta, n.now=n.now, i.now=i.now, s.now=s.now, r.now=r.now, b.now=b.now)

#Define Time Steps
T<- cumsum(rexp(length(simdata$I), rate=(1/2.2)))

#Plot Simulation
ggplot(simdata, aes(x=T)) +
  geom_line(aes(y=S, col="S"))+
  geom_line(aes(y=I, col="I"))+
  geom_line(aes(y=R, col="R"))+
  labs(x="Time Steps", y = "Number of Individuals", title = "Simulated Epidemic Dynamics")



```
```{r}
#Add Jitter to Infecteds 
S<- simdata$S
S_jitter <- jitter(S)

ggplot(simdata, aes(x=T)) +
  geom_line(aes(y=S_jitter, col="S_jitter"))+
  geom_line(aes(y=I, col="I"))+
  geom_line(aes(y=R, col="R"))+
  labs(x="Time Steps", y = "Number of Individuals", title = "Simulated Epidemic Dynamics")
```
```{r, echo=FALSE}
#Reconstructing Z dynamics 
cumbirths <- cumsum(simdata$B)
cuminfecteds <- cumsum(I)
cumulativeinfo <- as.data.frame(matrix(c(cumbirths, cuminfecteds), nrow=length(cuminfecteds)))
names(cumulativeinfo) <- c("cumbirths", "cuminfecteds")


#plot(cumbirths ~ cuminfecteds)

regression <- lm(cumbirths ~ 1, offset=(cuminfecteds))
#summary(regression)

Z <- as.vector(resid(regression))
#summary(Z)

ggplot(simdata, aes(x=T))+
  geom_line(aes(y=S_jitter, col="S_jitter"))+
  geom_line(aes(y=Z, col="Z"))+
  labs(x="Time", y="Number of Susceptibles", title="Reconstructed Dynamics")
```
```{r, include=FALSE}
#Preparing Data for Bayes
n <- simdata$N

time_added<-0
for(i in 2:(length(T))){
  time_added[i] <- T[i] - T[i-1]
}

data_z <- list(N=length(I), P=as.integer(n), I=as.integer(I), Z=as.integer(Z), time_added=time_added)

```
```{r, include=FALSE}
Stan_z <- as.character("
data {
	int<lower=1> N; // time steps
  real P[N]; //population size
  real time_added[N];
	int <lower=0> I[N]; // infecteds 
	int  Z[N]; // Zt 
}

parameters {
  real <lower=P[1]/100, upper= P[N]> Sbar;
	real<lower=0, upper=5> beta;
  real<lower=0, upper=5> gamma;
}

model {
	//priors
  Sbar ~ normal(P[1]/2, P[1]/10);
	beta ~ gamma(3, 1);
  gamma ~ gamma(2,1);

	//likelihood
	for (n in 2:N){
		I[n] ~ neg_binomial_2(beta*I[n-1]*(Z[n-1]+Sbar)/P[n-1], I[n-1]);
    time_added[n]~exponential_log(gamma);
  }
}

                            ")

bayes_reconstructed <- stan(model_code = Stan_z, data=data_z, iter=500000, chains=4)
```
```{r, echo=FALSE}
Sbar <- extract(bayes_reconstructed, 'Sbar')
Sbar <- unlist(Sbar, use.names=FALSE)
lowlim_Sbar <- quantile(Sbar, probs=.025, names=FALSE)
uplim_Sbar <- quantile(Sbar, probs=.975, names=FALSE)

S_lower <- lowlim_Sbar + Z
S_upper <- uplim_Sbar + Z

ggplot(simdata, aes(x=T))+
  geom_line(aes(y=S_jitter, col="S_jitter"))+
  geom_line(aes(y=S_lower, col="S_lower"))+
  geom_line(aes(y=S_upper, col="S_upper"))+
  labs(x="time", y="Number of Susceptibles", title="Reconstructed vs. Simulated Susceptible Dynamics")
```
```{r, include=FALSE}
##Running Bayes with True Susceptible Dynamics 
data_sim <- list(N=length(I), P=as.integer(n), I=as.integer(I), S=as.integer(S_jitter), time_added=time_added)

SIRstancode <- as.character("
                            data {
	int<lower=1> N; // time steps
  real P[N]; //population size
  real time_added[N];
	int <lower=0> I[N];
	int <lower=0> S[N];
}


parameters {
	real<lower=0, upper=5> beta;
  real<lower=0, upper=5> gamma;

}

model {
	//priors
	beta ~ gamma(3, 1);
  gamma ~ gamma(2,1);
	//likelihood
	for (n in 2:N){
		I[n] ~ neg_binomial_2(beta*I[n-1]*S[n-1]/P[n-1], I[n-1]);
    time_added[n]~exponential_log(gamma);
  }
}

                            ")

bayes_sim <- stan(model_code = SIRstancode, data=data_sim, iter=500000, chains=4)
```
```{r}
bayes_sim
bayes_reconstructed
```{r}

```


```{r}
#Initial Conditions
n.now <- 763000
s.now <- 762995
i.now <- n.now-s.now
r.now <- 0
beta <- 1.66
birthrate <- rgamma(1, shape=2, scale=.000215)
b.now <- 0





#Stan Code
Stan_reconstructed <- as.character("
data {
	int<lower=1> N; // time steps
  real P[N]; //population size
  real time_added[N];
	int <lower=0> I[N]; // infecteds 
	int  Z[N]; // Zt 
}

parameters {
  real <lower=P[1]/100, upper= P[N]> Sbar;
	real<lower=0, upper=5> beta;
  real<lower=0, upper=5> gamma;
}

model {
	//priors
  Sbar ~ normal(P[1]/2, P[1]/10);
	beta ~ gamma(3, 1);
  gamma ~ gamma(2,1);

	//likelihood
	for (n in 2:N){
		I[n] ~ neg_binomial_2(beta*I[n-1]*(Z[n-1]+Sbar)/P[n-1], I[n-1]);
    time_added[n]~exponential_log(gamma);
  }
}    ")


Stan_sim <- as.character("
                            data {
	int<lower=1> N; // time steps
  real P[N]; //population size
  real time_added[N];
	int <lower=0> I[N];
	int <lower=0> S[N];
}


parameters {
	real<lower=0, upper=5> beta;
  real<lower=0, upper=5> gamma;

}

model {
	//priors
	beta ~ gamma(3, 1);
  gamma ~ gamma(2,1);
	//likelihood
	for (n in 2:N){
		I[n] ~ neg_binomial_2(beta*I[n-1]*S[n-1]/P[n-1], I[n-1]);
    time_added[n]~exponential_log(gamma);
  }
}     ")
```

